---
title: "SPL: Fama French Factor Model"
author: "Xin Tan, Daria Skidnova, Jeff Giddens"
date: "2018/6/9"
output: beamer_presentation
bibliography: FamaFrench.bib
fontsize: 10pt
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
setwd("~/Dropbox/Study/R/Fama French/")
```


## Structure

\large

1. Model

2. Data 

3. Results

\normalsize

## 1. Model

- CAPM
$R_i-R_F = \beta\cdot(R_M-R_F)$
\newline

- @Fama1993 3 Factors
$R_i-R_F = \beta_M\cdot(R_M-R_F) + \beta_S\cdot SMB + \beta_V\cdot HML$
\newline

- @Fama2015 5 Factors
$R_i-R_F = \beta_M\cdot(R_M-R_F) + \beta_S\cdot SMB + \beta_V\cdot HML + \beta_P\cdot RMW + \beta_I\cdot CMA$


## 2. Data

- Model factors including the estimated market return and the risk-free rate from
Kenneth French's data library: http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html \newline

- S&P 500 stock prices downloaded with `BatchGetSymbols` library.

\small
```{r echo=TRUE, eval=FALSE}
library(BatchGetSymbols)
Companies <- GetSP500Stocks()
Stocks<- BatchGetSymbols(tickers = Companies$tickers, 
                     first.date = "2010-01-01", 
                     last.date = "2017-12-31")
```
\normalsize

## 2. Data: S&P 500

- `GetSP500Stocks()` returns company info like name and sector. \newline

- Downloaded stock data contains two dataframes: \newline
__df.control__ descriptive info. e.g. whether the download for the ticker is successful.\newline
__df.tickers__ downloaded price data, stacked in rows.\newline

- `BatchGetSymbols()` function __keeps__ stocks with 75% or more price points compared to the benchmark. \newline We took _3M Company_'s dates as they are complete.\newline

\small
```{r echo=TRUE, eval = FALSE}
good.tickers <- Stocks$df.control$
    ticker[Stocks$df.control$threshold.decision=="KEEP"]

SP500.data<-data.frame(date = Stocks$df.tickers$
             ref.date[Stocks$df.tickers$ticker=="MMM"]) 
```
\normalsize

## 2. Data: S&P 500

- Loop over the downloaded data __df.tickers__.
- Subset each company by its ticker.
- Rename `colname` of the price series to the ticker symbol.
- Merge to the existing dataframe by matching dates.

\small
```{r echo=TRUE, eval = FALSE}
for(i in 1:length(good.tickers))
{
  X <- data.frame(date = 
    Stocks$df.tickers$
      ref.date[Stocks$df.tickers$ticker==good.tickers[i]], 
    Stocks$df.tickers$
      price.adjusted[Stocks$df.tickers$ticker
                                ==good.tickers[i]])
  
  colnames(X)[2] <- Stocks$df.tickers$
        ticker[Stocks$df.tickers$ticker==good.tickers[i]]
  
  SP500.data <- merge.data.frame(SP500.data, X, 
                                by = "date", all.x = TRUE)
}
```
\normalsize


## 2. Data: from Daily Price to Monthly Return

__`quantmod::monthlyReturn()`__ requires non-`NA` daily prices in __`xts`__ format.

\small
```{r echo=TRUE, eval = FALSE}
Stock.Prices.Daily <- 
  xts(Stock.Prices.Daily[,-1], 
        order.by = as.POSIXct(Stock.Prices.Daily$date))

Stock.Prices.Daily <- 
  Stock.Prices.Daily[!is.na(Stock.Prices.Daily)]

Stock.Prices.Monthly <- monthlyReturn(Stock.Prices.Daily)
```
\normalsize

How to chose? 

- __Remove stocks with `NA`__ ensures remaining stocks have same number of observations. 

- __Remove `NA`s in each series__ results in a larger sample size.

Price data with `NA`s in the middle might result in strange monthly returns. 
\scriptsize("BHY" Brighthouse Financial Inc. removed for 2015-2017 runs)

## 2. Data: Linear Regression

- Linear Regression in R

```{r echo=FALSE}
library(readxl)
FF3<- read_excel("Data/FF3_196307-199112.xlsx")
# unlist: convert the data into vector format
rmrf<-unlist(FF3[,2])
P25<- read_excel("Data/FF3_25_ValueWeighted.xlsx")
rmrf<-unlist(FF3[,2])
smb<-unlist(FF3[,3])
hml<-unlist(FF3[,4])
rf<-unlist(FF3[,5])
ri<-unlist(P25[,2])
rirf<-unlist(ri-rf)
```

\small
```{r echo=TRUE}
y <- lm(rirf ~ rmrf + smb + hml);
round(summary(y)$coefficients, digits = 4)
```
\normalsize

- We can read out the regression summary from the `coefficients` matrix.\newline

- Similarly $R^2$ from `summary(y)$r.squared` \newline 
and standard errors from `summary(y)$sigma`.

## 3. Results

\large
Overview:
\normalsize

- Replicating @Fama1993

- Regression over S&P500 stocks

- 3 Factors vs. 5 Factors

- Stock selection?


## 3.1 Results: Replicating the 3 Factors Model
\noindent\makebox[\textwidth]{\includegraphics[width=\paperwidth]{FF3_Results.pdf}}

## 3.1 Results: Replicating the 3 Factors Model

Results from @Fama1993:
\noindent\makebox[\textwidth]{\includegraphics[width=\paperwidth]{FF1993-Table6.png}}


## 3.2 Regression over S&P 500 stocks (All)

\noindent\makebox[\textwidth]{\includegraphics[width=0.90\paperwidth]{Plots/SP500-1980-2015.png}}

## 3.2 Regression over S&P 500 stocks (Survivors)

\noindent\makebox[\textwidth]{\includegraphics[width=0.90\paperwidth]{Plots/SP500-1980-2015-168-stock.png}}


## 3.3 3 Factors vs. 5 Factors

- $R^2$ Comparison

\noindent\makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{Plots/R2-FF-SP500-2010-2017.png}}

\tiny\linespread{0.5}
CD: Consumer Discretionary, CS: Consumer Staples, E: Energy, FI: Financials, H: Health Care, I: Industrials,\newline 
IT: Information Technology, M: Materials, RE: Real Estate, T: Telecommunication Services, U: Utilities
\normalsize

## 3.3 3 Factors vs. 5 Factors

- P-Values: 3 Factors

\noindent\makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{Plots/P-FF3-SP500-2010-2017.png}}

\tiny\linespread{0.5}
CD: Consumer Discretionary, CS: Consumer Staples, E: Energy, FI: Financials, H: Health Care, I: Industrials,\newline 
IT: Information Technology, M: Materials, RE: Real Estate, T: Telecommunication Services, U: Utilities
\normalsize

## 3.3 3 Factors vs. 5 Factors

- P-Values: 5 Factors

\noindent\makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{Plots/P-FF5-SP500-2010-2017.png}}

\tiny\linespread{0.5}
CD: Consumer Discretionary, CS: Consumer Staples, E: Energy, FI: Financials, H: Health Care, I: Industrials,\newline 
IT: Information Technology, M: Materials, RE: Real Estate, T: Telecommunication Services, U: Utilities
\normalsize



## 3.3 3 Factors vs. 5 Factors

- Regressed Coefficients: 3 Factors

\noindent\makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{Plots/b-FF3-SP500-2010-2017.png}}

\tiny\linespread{0.5}
CD: Consumer Discretionary, CS: Consumer Staples, E: Energy, FI: Financials, H: Health Care, I: Industrials,\newline 
IT: Information Technology, M: Materials, RE: Real Estate, T: Telecommunication Services, U: Utilities
\normalsize

## 3.3 3 Factors vs. 5 Factors

- Regressed Coefficients: 5 Factors

\noindent\makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{Plots/b-FF5-SP500-2010-2017.png}}

\tiny\linespread{0.5}
CD: Consumer Discretionary, CS: Consumer Staples, E: Energy, FI: Financials, H: Health Care, I: Industrials,\newline 
IT: Information Technology, M: Materials, RE: Real Estate, T: Telecommunication Services, U: Utilities
\normalsize


## 3.4 Stock Selection?

- Stock returns from Jan. 2010 to Dec. 2017. \newline 
Corr. plot of the stock returns and the regressed coefficients.

- _alpha_ or _Intercept_ highly correlated to stock returns.

- Positive corr. to SMB: Small cap premium.\newline
Negative corr. to HML: Market seemed to favor __Growth__ stocks (instead of __Value__) in the bullish years.

- Negative corr. to RMW and CMA.

\noindent\makebox[\textwidth]{
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF3-all-2010-2017.png}
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF5-all-2010-2017.png}
}

## 3.4 Stock Selection: 3 Factor / Top and Bottom 20s

- Winners have positive correlations to both _SMB_ and _HML_, \newline while losers have negative correlations.

- Small Cap and Value premiums work at the extremes.

\noindent\makebox[\textwidth]{
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF3-Top20-2010-2017.png}
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF3-Bottom20-2010-2017.png}
}

Although the overall winner, _NETFLIX_, is clearly neither _small cap_ nor _value_.

## 3.4 Stock Selection: 5 Factor / Top and Bottom 20s

- __Profitability__ factor _RMW_: negatively corr. to returns of winners and positively corr. to that of losers: might not be good indicator for stock selection. \newline Negative regressed beta for _RMW_ for certain sectors: companies might have poor near-term operating performance but very high future potentials, which is _priced-in_ to the current stock prices.

- __Investment__ factor _CMA_ is negatively correlated to the returns of both winners and losers: both winners and losers invest aggressively? 



\noindent\makebox[\textwidth]{
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF5-Top20-2010-2017.png}
  \includegraphics[width=0.45\paperwidth]{Plots/Corrplots/SP500-FF5-Bottom20-2010-2017.png}
}

## 3.4 Stock Selection: 5 Factor / Top and Bottom Factors

- Calculate return percentile for each stock (best return ~ 1, worst ~ 0).

- Filter out largest and smallest estimated coefficient for each factor.

- With and without considering the significance level of 5%.

\noindent\makebox[\textwidth]{\includegraphics[width=\paperwidth]{Plots/Return-Percentile-2010-2017.png}}

- Not considering significance, market seems to favor __growth__ stocks.

- Most of the stocks with lowest estimated coefficient for _HML_ belongs to _Health Care_ or _Consumer Discretionary_. 

## Thank You!

\large Below are backup slides
\normalsize

## Melt and Plotting 

- Regression results in dataframe `Results`. 

- `reshape::melt()` prepares the data for plots.

- Each Fama French factor will have 1 sub-plot comparing different sectors \newline

\scriptsize
```{r echo=TRUE, eval = FALSE}
df.melt <- melt(
  Results[,c("Intercept", "Mkt-Rf", "SMB", "HML", "RMW", "CMA", "Sec")], 
  "Sec"
)
```
\normalsize

- Boxplot with the melted data

\scriptsize
```{r echo=TRUE, eval = FALSE}
ggplot(df.melt, aes(x=Sec, y=value)) 
  + geom_boxplot() 
  + facet_wrap(~ variable, scales='free')
```
\normalsize

## Melt and Plotting

- Illustration of `reshape::melt()`: \newline
Data is reshaped into one dimensional vector with group / subgroups in additional columns.

\scriptsize
```{r echo=TRUE, eval = FALSE}
df.melt <- melt(
  Results[,c("Intercept", "Mkt-Rf", "SMB", "HML", "RMW", "CMA", "Sec")], 
  "Sec"
)
```
\normalsize

\noindent\makebox[\textwidth]{\includegraphics[width=0.8\paperwidth]{melt_example.pdf}}


## Helper functions

- Resizing the regression output from an 1 dimensional vector to the same format as in the papers with rows for _SMB_ and columns for _HML_.

\scriptsize
```{r echo=TRUE, eval=FALSE}
resize <- function(x)
{
  return(
    matrix(x, nrow=5, byrow = TRUE, 
        dimnames = list(
          c("SMALL", "2", "3", "4", "BIG"), 
          c("LOW", "2", "3", "4", "HIGH")
        )
    )
  )
}
```
\normalsize

## Descriptions copied from K.French's Website

Description of the factors:
<http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/Data_Library/f-f_5_factors_2x3.html>

- SMB (Small Minus Big) is the average return on the nine small stock portfolios minus the average return on the nine big stock portfolios.
- HML (High Minus Low) is the average return on the two value portfolios minus the average return on the two growth portfolios. 

- RMW (Robust Minus Weak) is the average return on the two robust operating profitability portfolios minus the average return on the two weak operating profitability portfolios.
- MA (Conservative Minus Aggressive) is the average return on the two conservative investment portfolios minus the average return on the two aggressive investment portfolios.


## Descriptions copied from K.French's Website (cont.)

- Rm-Rf, the excess return on the market, value-weight return of all CRSP firms incorporated in the US and listed on the NYSE, AMEX, or NASDAQ that have a CRSP share code of 10 or 11 at the beginning of month t, good shares and price data at the beginning of t, and good return data for t minus the **one-month Treasury bill rate** (from Ibbotson Associates).

## References
