---
title: "SPL Fama French"
bibliography: FamaFrench.bib
output:
  pdf_document: default
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Structure

1. Introduction

2. Data Preparation

3. Simple Regression

4. Replicating the 3-Factor model

5. S&P500 Results

6. Going 5-Factor



# 1. Introduction

The Fama French model is a model for explaining stock returns. It extends the classical Capital Asset Pricing Model (CAPM) by having additional factors. 

$$R_i-R_F = \beta\cdot(R_M-R_F)$$

@Fama1993 introduces *SMB* (Small market cap Minus Big / Size) and *HML* (High book-to-market Minus Low / Value) to capture the observation that small capitalization and high book value to market value ("value" in contrast to "growth") stocks tend to outperform the market. 

$$R_i-R_F = \beta_M\cdot(R_M-R_F) + \beta_S\cdot SMB + \beta_V\cdot HML$$

@Fama2015 adds *RMW* (Robust operating profit Minus Weak / Profitability) and *CMA* (Conservative investment strategy Minus Aggressive / Investment). 

$$R_i-R_F = \beta_M\cdot(R_M-R_F) + \beta_S\cdot SMB + \beta_V\cdot HML + \beta_P\cdot RMW + \beta_I\cdot CMA$$

Fama French factors are calculated as return spreads between two portfolios, e.g. SMB is the difference between the return of a small cap portfolio and that of a large cap portfolio.

We choose the Fama French model due to the high quality data available at [Kenneth R. French's data library](http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html)

Refer to [Wikepedia](https://en.wikipedia.org/wiki/Fama%E2%80%93French_three-factor_model) for more information.

\newpage

# 2. Data Preparation

## 2.1 Fama French Data

[French's data library](http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html) contains data for the factors, corresponding market returns and risk free rates, as well as the portfolios returns featured in the papers:

* **3 Factors** 1926.07.01 to 2018.03.29 as daily / weekly / monthly data

* **5 Factors** 1963.07.01 to 2018.03.29 as daily / monthly / yearly data

* **25 Portfolios (5x5)** formed on Size and Book-to-Market 1926.07 to 2018.03 corresponding to the @Fama1993 3-factor setup (P24 Table 6). 

The downloaded CSV data contains headers and footers that need to be removed before input to R.

## 2.2 S&P 500 Stock Data

The [`BatchGetSymbols`](https://cran.r-project.org/web/packages/BatchGetSymbols/index.html) library has a function `BatchGetSymbols()` for downloading S&P500 stock prices and volumes from a cached repository, thus avoiding problems when downloading large amount of data directly from Yahoo or Google (e.g. the `getSymbols` function from the [`quantmod`](https://www.quantmod.com/documentation/getSymbols.html) library) 


```{r echo=FALSE, results='hide', message=F, warning=F}
# required by BatchGetSymbols
library(xml2)
library(rvest)
```

```{r echo=TRUE, results='hide', message=F, warning=F}
library(BatchGetSymbols)

# Get company information incl. tickers for SP500 stocks
Companies <- GetSP500Stocks()
```

```{r echo = T}
kable(head(Companies, n=5)[,1:5])
kable(head(Companies[,6:9], n=5))
```


Function `GetSP500Stocks()` returns S&P500 company information including name, tickers and sectors. For downloading the price data, we only need the tickers.


```{r echo=TRUE, results='hide', message=F, warning=F}
# Batch download data from Yahoo Finance
Stocks<- BatchGetSymbols(tickers = Companies$tickers, 
                     first.date = "2017-01-01", 
                     last.date = "2017-12-31")
```


The downloaded list contains 2 dataframes:

* **df.control** contains descriptive information like whether the download for the ticker is successful.

* **df.tickers** contains the downloaded price data. Each row is the price data for one ticker at one date, hence we need to process the data into a format easier to work with. 


(Use `kable()` function in `Knitr` library to format table output in PDF.)


```{r echo = T}
kable(head(Stocks$df.control, n=3)) 
kable(head(Stocks$df.tickers[,1:5], n=3)) 
kable(head(Stocks$df.tickers[,6:10], n=3))
```


Below code selects the downloaded tickers (marked by `df.control$threshold.decision=="KEEP"`) and use the dates from 3M as the date column for dataframe `SP500.data`. 


It reads stocks ticker by ticker and matches previous price series by date. The unmatched dates will have `NA`s. The new stock price series is merged into the dataframe as a new column with the ticker symbol as the column name.


```{r eval = FALSE}
good.tickers <- Stocks$df.control$
        ticker[Stocks$df.control$threshold.decision=="KEEP"]

# Fill dates as the first stock "MMM" happens to have complete dates 
# (column name = "date")
SP500.data<-data.frame(date = Stocks$
                         df.tickers$
                         ref.date[1:max(Stocks$df.control$total.obs)]) 

for(i in 1:length(good.tickers))
{
  # X is a temp dataframe that has 2 columns, 
  # 1st is date (for matching), 2nd is the actual data (e.g. closing price)
  
  # Choose relevant data by matching tickers
  X <- data.frame(date = 
        Stocks$df.tickers$ref.date[Stocks$df.tickers$ticker==good.tickers[i]], 
        Stocks$df.tickers$price.adjusted[Stocks$df.tickers$ticker==good.tickers[i]])
  
  # change the column name of X to be the ticker of the stock
  # colnames(X)[2] = good.tickers[i] # this one don't work
  colnames(X)[2] <- Stocks$df.tickers$
        ticker[Stocks$df.tickers$ticker==good.tickers[i]]
  
  # merge X as a new column into SP500.data by matching date
  # missing dates will have NA by default
  SP500.data <- merge.data.frame(SP500.data, X, by = "date", all.x = TRUE)
}
```


We write the processed data to CSVs.

\newpage

# 3. Simple Regression

```{r echo=FALSE}
setwd("~/Dropbox/Study/R/Fama French")
```

`readxl` library for reading Excel data. 

The imported data would be stored as `data.frame` and must be `unlist()` into vectors for regression.
(data.frame is also a list in R)


```{r}
library(readxl)
FF3<- read_excel("Data/FF3_196307-199112.xlsx")
# unlist: convert the data into vector format
rmrf<-unlist(FF3[,2])
```

```{r echo=FALSE}
P25<- read_excel("Data/FF3_25_ValueWeighted.xlsx")
rmrf<-unlist(FF3[,2])
smb<-unlist(FF3[,3])
hml<-unlist(FF3[,4])
rf<-unlist(FF3[,5])
ri<-unlist(P25[,2])
rirf<-unlist(ri-rf)
```


OLS regression can be performed with two lines of code:


```{r}
y <- lm(rirf ~ rmrf + smb + hml);
summary(y)
```

`summary(y)` contains the regression results and specific results could be obtained, e.g., via:

```{r eval=FALSE}
summary(y)$coefficients
```

which returns the regression betas and their standard errors, t-values and p-values in a matrix.

```{r kable,echo=FALSE}
kable(summary(y)$coefficients, digits=2)
```

\newpage
# 4. Replicating the 3-Factor model

To check that we have implemented the Fama French model correctly, we try to replicate the results of table 6 of @Fama1993 which involves monthly return data of 25 value-weighted portfolios from July 1963 to December 1991.

The data set structures the data as 1 column of months (*YYYYDD* format) plus 25 columns of portfolio monthly returns. The first return column is SMALL (market cap) LoBM (low book-to-market / "growth"). The first 5 return columns are all small cap but with increasing book-to-market ratios. The last 5 return columns are all large cap with the last column being BIG (market cap) HiBM (high book-to-market / "value").

In reporting, results are structured in a matrix with rows representing market cap and columns for book to market ratios.

## 4.1 Batch regression

With the OLS regression code working, below code runs regression on each portfolio and saves the results in a list `results`. 

```{r}
# Store summaries into a results list
results <- list()
# The first column of P25 is dates, not data
for(i in 1:(ncol(P25)-1))
{
  rirf<-unlist(P25[,i+1])-rf # Data starts from the 2nd col of P25
  y<-lm(rirf~rmrf+smb+hml)
  results[[i]]<-summary(y)
}
```

## 4.2 Formatting the results

We then read out the results, stack them into corresponding vectors, then reshape them into the $5\times 5$ format as in the paper for ease of comparison. Results are highly similar and we have not yet identify why they do not match exactly, perhaps due to rounding errors.

```{r}
betas <- vector()
std.errors <- vector()
t.values <- vector()
R.squareds <- vector()
# save all betas 
for(i in 1:(ncol(P25)-1))
{
  betas <- cbind(betas,results[[i]]$coefficients[,1])
  std.errors <- cbind(std.errors,results[[i]]$sigma)
  t.values <- cbind(t.values, results[[i]]$coefficients[,3])
  R.squareds <- cbind(R.squareds, results[[i]]$adj.r.squared)
}

# resize the output to 5x5 format like Fama French paper
resize <- function(x)
{
  df = data.frame(matrix(x, nrow=5, byrow = TRUE))
  colnames(df) = c("Low", "2", "3", "4", "High")
  rownames(df) = c("Small", "2", "3", "4", "Big")
  return(df)
}
# resize alpha
alpha <- resize(betas[1,])
kable(alpha, digits=2)

# resize beta
market.beta <- resize(betas[2,])
SMB.beta <- resize(betas[3,])
HML.beta <- resize(betas[4,])

# display beta below

kable(market.beta, digits=2)
kable(SMB.beta, digits=2)
kable(HML.beta, digits=2)
```


\includegraphics{FF1993-Table6.PNG}


Similarly for t-statistics and $R^2$: 


```{r}
# resize t-stats
market.t <-resize(t.values[2,])
SMB.t <- resize(t.values[3,])
HML.t <- resize(t.values[4,])

kable(market.t, digits=2)
kable(SMB.t, digits=2)
```

\newpage
```{r}
kable(HML.t, digits=2)

# resize R-squareds
kable(resize(R.squareds), digits=2)
kable(resize(std.errors), digits=2)
```

\includegraphics{FF1993-Table6-2.PNG}

\newpage
# 5. S&P500 Results

We first apply the above methods on the downloaded S&P 500 stocks' price returns to see if there is any pattern with the regression results. Also to test out the code for handling hundreds of stocks.

Then we separate the data by 5-year periods and loop over both years and stocks to see if patterns change over time. 

## 5.1 Running the model for S&P 500 stocks

Below code works as follows:

1. Read-in price data and do the necessary formatting.

2. Frame the data to the desired time period.

3. Convert the price data series into XTS series as required by 5.

4. Remove stocks with `NA`s in the series. 

5. Use `quantmod` library's `monthlyReturn()` function to batch convert the whole price matrix into a monthly return matrix. 

We need to remove `NA`s for using the `monthlyReturn()` function. Most `NA`s are due to data not available on the starting date of the series, e.g. the company has not IPO yet.

Here we face choices:

* Remove all columns with `NA`s, then all remaining stocks could have the regression in the same period, i.e. with the same number of observations. (This section)

* Dynamically frame the data based on the available non-`NA` data points, but then some stocks in the regression analysis will have fewer observations. (Tested in Section 5.2)



```{r message=F, warning=F}
library(quantmod)

# Read SP500 daily data and convert date column to date format
SP500.data <- read.csv("Data/SP500_price.adjusted_2010-2017.csv")
SP500.data$date <- as.Date(SP500.data$date)

# Select 2010 - 2017 range
Stock.Prices.Daily <- SP500.data[SP500.data$date>="2010-01-01" & 
                                   SP500.data$date<="2017-12-31",-1]

# Current FF3 till 201803, monthly
FF3 <- read.csv("Data/original/FF3.csv")
FF <- FF3[FF3$X >= 201001 & FF3$X <= 201712,]

# Convert series to XTS for using quantmod's monthlyReturn function
Stock.Prices.Daily <- xts(Stock.Prices.Daily[,-1], 
                          order.by = as.POSIXct(Stock.Prices.Daily$date))

# Number of stocks to start with
ncol(Stock.Prices.Daily)

# Remove stocks with NAs in the series, otherwise monthly Return will not work properly
Stock.Prices.Daily <- Stock.Prices.Daily[,colSums(is.na(Stock.Prices.Daily)) == 0]

# Apply monthlyReturn function to each column (it seems it converts only one column at a time)
Stock.Prices.Monthly <- do.call(cbind, lapply(Stock.Prices.Daily, monthlyReturn))
# Stock.Prices.Monthly <- na.omit(Stock.Prices.Monthly)
colnames(Stock.Prices.Monthly) <- colnames(Stock.Prices.Daily)

# Number of stocks left
ncol(Stock.Prices.Monthly)

```

As in this example, we start with 465 stocks and remove 23 stocks with incomplete data.

Then the regression part is similar to Section 4.1, except that we need to transpose the coefficients to get the dimensions right before stacking them together column by column, with each column representing one stock.


```{r}
Results <- list()
for(i in 1:ncol(Stock.Prices.Monthly))
{
  RiRF <- Stock.Prices.Monthly[,i] - FF$RF
  Regression <- lm(RiRF ~ FF$Mkt.RF + FF$SMB + FF$HML)
  Results[[i]] <- summary(Regression)
}

# Results!
betas <- vector()
std.errors <- vector()
t.values <- vector()
p.values <- vector()
r.squareds <- vector()
adj.r.squareds <- vector()

for(i in 1:ncol(Stock.Prices.Monthly))
{
  betas <- cbind(betas,Results[[i]]$coefficients[,1])
  std.errors <- cbind(std.errors,Results[[i]]$sigma)
  t.values <- cbind(t.values, Results[[i]]$coefficients[,3])
  p.values <- cbind(p.values, Results[[i]]$coefficients[,4])
  
  r.squareds <- cbind(r.squareds, Results[[i]]$r.squared)
  adj.r.squareds <- cbind(adj.r.squareds, Results[[i]]$adj.r.squared)
  
}

Regression.results <- cbind(data.frame(colnames(Stock.Prices.Monthly)), 
                    t(r.squareds), t(adj.r.squareds), 
                    t(betas), t(p.values))

colnames(Regression.results) = c("Ticker", "R.Squared", "Adj.R.Squared", 
                         "Intercept", "Mkt.Rf", "SMB", "HML", 
                         "P(Intercept)", "P(Mkt.Rf)", "P(SMB)", "P(HML)")

```


We add company information like name and sector to make the results easier to understand. The constituent data is from a downloaded [CSV file](https://github.com/datasets/s-and-p-500-companies), which can also be found in the downloaded data introduced in Section 2.2.

We use a left join (`merge()` function with parameter all.x = TRUE) to add company name and sector to our regression results.


```{r}
# Read in SP500 company ticker information
Mapping <- read.csv("Data/constituents.csv")
colnames(Mapping)[1] <- "Ticker"
Regression.results <- merge(x = Regression.results, y = Mapping, by = "Ticker", all.x = TRUE)
```

Then we can easily filter out specific companies, e.g. companies and sectors whose returns have the highest $Rˆ2$ in the Fama French model. Interesting to see Financials come on top:


```{r}
# select stocks with R2>=0.08
R2 <- Regression.results[Regression.results$R.Squared>=0.08,
                   c("Ticker","Name","Sector","R.Squared")]

# sort with R2 from largest to smallest, get top 10
kable(head(R2[order(R2$R.Squared, decreasing = T),], n=10), digits = 4)
```

We could also box-plot the distribution of the betas and their p-values. A new column is needed for using the `melt()` function (`reshape2` library) for the convenience of box-plot. In general, each column in the dataframe will be plotted into a separated graph, while data within each column is grouped by the value in the added column. 
Hence in the below code, the original data frame contains two columns: the estimated $\beta$'s and their p-values. The added column in the dataframe marks which rows are the estimated coefficients for intercept, which rows are the estimated $\beta_M$, etc.


```{r}
# boxplot of regression results
library(ggplot2)
library(reshape2)
num.stocks <- dim(Regression.results)[1]

plot.data <- data.frame(Betas = rep(c("Intercept", "Mkt-Rf", "SMB", "HML"),
                        rep(num.stocks, 4)))

plot.data$Level <- as.vector(cbind(
                    Regression.results$Intercept, 
                    Regression.results$Mkt.Rf, 
                    Regression.results$SMB, 
                    Regression.results$HML))

plot.data$P.Value<- as.vector(cbind(
                    Regression.results$`P(Intercept)`, 
                    Regression.results$`P(Mkt.Rf)`,
                    Regression.results$`P(SMB)`, 
                    Regression.results$`P(HML)`))

plot.melt <- melt(plot.data, "Betas")
ggplot(plot.melt, aes(x=Betas, y=value)) + geom_boxplot() + 
                    facet_wrap(~ variable, scales='free')
```

From the p-values, *SMB* and *HML* are not significant for many stocks. 


## 5.2 Running the model for each 5-year period from 1980 to 2015

Data downloaded with `BatchGetSymbols` has an issue that the earlier the series (e.g. in the 1980s), the less stocks are available, most probably due to stocks being replaced in the S&P 500 index. To fix this issue, we could either:

1. Get the constituents for S&P 500 for each period and download those exact tickers, which may not work due to data availability. Even if it worked, we might be comparing apples to oranges, if the set of companies change over time.

2. Limit the data set to companies that survive over time. But then we have a much smaller set and miss out large names like Google or Facebook since they IPO in the 2000s.

Currently we simply take all the data available for each period for the regression, thus the results should be interpreted with a grain of salt.

Code is built based on Section 5.1, except that we stored only the results needed for plotting. Here in the document the `print()` and `cat()` functions are muted as they were merely for displaying the progress of the code in run time. Library `lubridate` provides some nice functions like `year()` for handling dates.


```{r results='hide'}
# loop over above codes to regress data from 1980 - 2015, group every 5 yrs.
library(lubridate)
```
```{r}
List.of.start.date <- seq(as.Date("1980/1/1"), as.Date("2016/1/1"), "years")
List.of.start.date <- List.of.start.date[year(List.of.start.date)%%5==0]

# FF3: 192607 - 201803, monthly
FF3 <- read.csv("Data/original/FF3.csv")

# Each batch stores results for a 5yr group
Batch <- list()
Descriptions <- list()

Beta.batch <- list()

for(i in 1:(length(List.of.start.date)-1))
{
  start.date <- as.Date(List.of.start.date[i])
  end.date <- as.Date(List.of.start.date[i+1])-1
  # print(paste(start.date, end.date,sep=" - "))
  
  # read data
  file.name <- paste("Data/SP500_price.adjusted_", 
                     paste(year(start.date), year(end.date), sep="-"), ".csv", sep="")
  SP500.data <- read.csv(file.name)
  SP500.data$date <- as.Date(SP500.data$date)
  
  # remove first column "X" created due to importing  
  Stock.Prices.Daily <- SP500.data[SP500.data$date>= start.date & 
                                     SP500.data$date<= end.date,-1]
  
  # Convert series to XTS for using quantmod's monthlyReturn function
  Stock.Prices.Daily <- xts(Stock.Prices.Daily[,-1], 
                            order.by = as.POSIXct(Stock.Prices.Daily$date))
  
  # try a diff approach: loop over stocks and convert to monthly for each stock
  
  # initialize
  Results <- list()
  Description <- data.frame()
  
  betas <- data.frame()
  
  # loop through stocks
  for(j in 1:ncol(Stock.Prices.Daily))
  {
    # The j-th stock
    Rj <- Stock.Prices.Daily[,j]
    
    # cat(colnames(Stock.Prices.Daily[,j]), " ")
    # non-NA entries
    Rj <- Rj[!is.na(Rj),]
    Rj <- monthlyReturn(Rj)
    
    # matching FF data
    FF <- FF3[FF3$X >= format(index(head(Rj, n=1)), "%Y%m") & 
              FF3$X <= format(index(tail(Rj, n=1)), "%Y%m"), ]
    
    # Rj is now RjRF
    Rj <- Rj-FF$RF
    Regression <- lm(Rj ~ FF$Mkt.RF + FF$SMB + FF$HML)
    Results[[j]] <- summary(Regression)
    Description <- rbind(Description, 
                         data.frame(colnames(Stock.Prices.Daily[,j]), 
                                    format(index(head(Rj, n=1)), "%Y%m"), 
                                    format(index(tail(Rj, n=1)), "%Y%m"), 
                                    length(Rj)))
    
    # try read-out results at regression time
    # betas, p-values, r-squareds
    betas <- rbind(betas, cbind(data.frame(t(Results[[j]]$coefficients[,1])),
                                data.frame(t(Results[[j]]$coefficients[,4])),
                                data.frame(t(Results[[j]]$r.squared))))
  }
  
  # Save all regression summaries
  Batch[[i]] <- Results
  
  # Save the ticker / dates for ease of tracking the regression summary
  colnames(Description) = c("Ticker", "Start.Month", "End.Month", "Number.of.Months")
  Descriptions[[i]] <- Description
  
  # Save the regression results for plotting
  colnames(betas) <- c("Intercept", "Mkt-Rf", "SMB", "HML", 
                       "P(Intercept)", "P(Mkt-Rf)", "P(SMB)", "P(HML)",
                       "R-squared")
  
  # Try rbind here instead of list for convenience of melt.
  Beta.batch[[i]] <- betas
  
  # remove temp variables
  rm(Description, Results, Regression, Rj, betas)
}

```


Similar to Section 5.1, we use `melt()` function and `ggplot()` for visualizing the results:


```{r}
df <- data.frame()
Num.Obs <- data.frame()
for(i in 1:(length(List.of.start.date)-1))
{
  start.date <- as.Date(List.of.start.date[i])
  end.date <- as.Date(List.of.start.date[i+1])-1
  
  label <- paste(year(start.date), year(end.date),sep="-")
  df <- rbind(df, cbind(rep(label, dim(Beta.batch[[i]])[1]), Beta.batch[[i]]))
  Num.Obs <- rbind(Num.Obs, 
                   cbind(  paste(year(start.date), year(end.date),sep="-"), 
                           dim(Beta.batch[[i]])[1]))
}

colnames(df) <- c("Year",
                  "Intercept", "Mkt-Rf", "SMB", "HML", 
                  "P(Intercept)", "P(Mkt-Rf)", "P(SMB)", "P(HML)",
                  "R-squared")

```
```{r eval=FALSE}
df.melt <- melt(df, "Year")
ggplot(df.melt, aes(x=Year, y=value)) + geom_boxplot() 
              + facet_wrap(~ variable, scales='free')
```


Regressing the data in different periods tells that the regression coefficients have changed over time. The explanatory power of the Fama French does not stay constant. Interestingly during 1990 to 1994 when @Fama1993 was published, *SMB* is most significant from p-values. 


\includegraphics{Plots/SP500-1980-2015.PNG}


\newpage
As emphasized earlier, the results are probably due to having varying stocks in each period:


```{r echo=FALSE}
colnames(Num.Obs) <- c("Time Period", "Number of Stocks")
kable(Num.Obs)
```


If we filter out stocks surviving all periods, we get 168 tickers. 
Surprisingly, results has only minor changes.


\includegraphics{Plots/SP500-1980-2015-168-stock.PNG}

\newpage
## 5.3 Stock Selection


A natural question arised from the study is whether we could use the model for stock selection. We can easily calculate period returns by $R_i = P_T / P_0 -1$ from the first and last adjusted closing prices of each stock.


The top and bottom 10 stocks in terms of gross returns from January 2010 to December 2017 are:


```{r echo = F}
Top20<-read.csv("FF3-Top20-2010-2017.csv")
colnames(Top20)<-c("X","Ticker","Ri","Name","Sector","Sec","R2","Adj.R2","a","Mkt","SMB","HML","P(a)","P(M)","P(S)","P(H)")
kable(Top20[1:10,c(2,4,6,3,7,9:16)], digits = 2)
Bottom20<-read.csv("FF3-Bottom20-2010-2017.csv")
colnames(Bottom20)<-c("X","Ticker","Ri","Name","Sector","Sec","R2","Adj.R2","a","Mkt","SMB","HML","P(a)","P(M)","P(S)","P(H)")
kable(Bottom20[1:10,c(2,4,6,3,7,9:16)], digits = 2)
```


We can use the `cor()` function to calculate the correlation matrix of data series and the `corrplot` library for plotting. 


**All Stocks / Top 20 / Bottom 20**


\includegraphics{Plots/SP500-FF3-2010-2017.png}


Plotting the correlations between regression results and stock returns reveal no particular pattern except for the intercept term in general. Top 20 stock returns do show positive correlations of the 3 factors, while the bottom 20 show negative correlations, which is consistent with the rationale behind the factors.



# 6. Going 5-Factor


@Fama2015 adds two additional factors *RMW* and *CMA*:

* **RMW**: Profitability factor: the return of **R**obust (profitability) stocks **M**inus **W**eak ones.

* **CMA**: Investment factor: the return of **C**onservative (low investment) firms **M**inus the **A**ggressive (high investment) ones.


The process is mostly identical to section 5.1 except for adding the two factors into regression. We tested on 2010-2017 data and identify a data issue with the downloaded S&P500 data: Ticker "BHY" *Brighthouse Financial Inc.* which has a large gap of `NA`s in 2016. It was not revealed in section 5, as 5.1 removed all stocks with `NA`s while 5.2 was tested with 1980 to 2015 data.

We have the following code to address this problem in the beginning, but decided to drop the BHY due to seemingly wrong results. Hence the actual code only needs to handle `NA`s at the beginning and at the end of the series, but not abnormalities in between.

```{r eval=FALSE}
# Pick out non-NA entries and convert to monthly return
Ri <- Stock.Prices.Daily[, i]
Ri <- Ri[!is.na(Ri),]
Ri <- monthlyReturn(Ri)

# Convert the existing row index to YYYYMM format to match the Fama French data
Ri <- data.frame(date = format(index(Ri), "%Y%m"), Ri)

# Select the matching FF periods
# Actually we do not need this one as 
# the next code will match the relevant periods anyway
FF <- FF5[FF5$X >= format(index(head(Ri, n=1)), "%Y%m") & 
          FF5$X <= format(index(tail(Ri, n=1)), "%Y%m"), ]
  
# New matching: dropped and revert to old matching because we exclude BHY
FF <- FF5[Ri$date,]

# Change due to Ri is now dataframe with two columns (date, return)
RiRF <- Ri$monthly.returns - FF$RF

```

We create short codes for sectors for ease of plotting.

```{r echo=FALSE}
Mapping<-read.csv("Data/constituents.csv")
T1 <- data.frame(table(Mapping$Sec))
colnames(T1) <- c("Sec", "Number of Companies")
T2 <- unique(Mapping[,c("Sector","Sec")])
T2 <- merge(x = T2, y = T1, by = "Sec")
# T2 <- T2[order(-T2[,3]),]
kable(T2)
rm(T1, T2)
```

Results show a large jump in $R^2$ for Telecommunication sector but then it contains only 3 companies

```{r echo=FALSE}
kable(Mapping[Mapping$Sec == "T",], row.names = F)
```


\newpage
$Rˆ2$ comparison: Fama French 3 Factors vs. 5 Factors. 

$Rˆ2$'s are generally higher with the 5 Factor models. 

\includegraphics{Plots/R2-FF-SP500-2010-2017.png}


The most noticeable difference between 3 factors and 5 factors is also with Telecommunication sector, with p-values for *HML* (value) and *CMA* (investment) being much smaller (significant) in the 5-factor model.

\newpage
Fama French 3 Factors:

\includegraphics{Plots/P-FF3-SP500-2010-2017.png}


Fama French 5 Factors:

\includegraphics{Plots/P-FF5-SP500-2010-2017.png}


\newpage
The regressed coefficients from the 3 Factors model:

\includegraphics{Plots/b-FF3-SP500-2010-2017.png}


and 5 Factors model:

\includegraphics{Plots/b-FF5-SP500-2010-2017.png}



We can use the same method as section 5.3 to visualize the correlation between stock returns and the regression results of the Fama French 5 Factors model. Here the top 20 are the 20 stocks with the greatest returns from January 2010 to December 2017, same as in section 5.3. 


**All Stocks / Top 20 / Bottom 20**


\includegraphics{Plots/SP500-FF5-2010-2017.png}


A large portion of the return is still captured by *alpha* the intercept.
Surprisingly, the top 20 stock returns show negative correlations with the added *RMW* and *CMA* factor, while the bottom 20 stock returns show positive correlations with *RMW* and still negative correlations with *CMA*. This proves that one cannot predict the future from the past: 


1. Winning stocks might not have robust operating profitability: a large portion of earnings is invested.


2. Winning stocks seem to benefit from past aggressive investments. \newline




Another perspective is we could look at the regressed factor values and see whether we can select stock based on these values. We first calculate the percentile for each stock return using the `ecdf()` function: below code first defines our percentile function by supplying the all stock returns as a vector, then the `ecdf_percentile()` function can return a vector of percentiles, given a vector of returns.


```{r echo=TRUE, eval=FALSE}
# Define function
ecdf_percentile <- ecdf(Results$Return)
# Apply function. 
ecdf_percentile(Results$Return)
```


Among 443 stocks with regression results, we take top 20 and bottom 20 stocks for the estimated coefficient of each factor. We then boxplot their return percentiles. Stock with greatest return from 2010 to 2017 will have a return percentile close to 1, stocks with poor returns will have a percentile close to 0.


We perform the selection with and without filtering for significance of the estimated coefficient.


\includegraphics{Plots/Return-Percentile-2010-2017.png}


From the results, if we only consider value without significance, *HML* shows strong separation power that **Growth** stocks with low book-to-market ratio outperform in this period, while **Value** stocks perform below average. Looking into the 20 stocks with the lowest estimated *HML* coefficient reveals that they are mostly Health Care or Comsumer Discretionary. When limiting scope to coefficients that are significant at 5% level, however, shows no particular separation power of all Fama French factors. Interestingly, at 5% significance, stocks with largest market exposure and smallest market exposure (CAPM beta / beta for Mkt-Rf) all perform below average.


```{r echo=FALSE}
temp <- read.csv("FF5-Return-Factor-Bottom-Not-Significant.csv")
temp <- temp[,-1]
temp <- data.frame(temp[temp$Factor=="HML", c("Ticker", "Name", "Sector", "Factor", "Factor.Value", "Factor.P", "Return.2010.2017", "Return.Percentile")])
colnames(temp)<-c("Ticker", "Name", "Sector", "Factor", "Est.Beta", "P-Value", "Return", "Percentile")
kable(temp, row.names = FALSE, digits = 2) 
```

# References
